---
output:
  pdf_document: default
  html_document: default
---


```{r, error=TRUE}
# Load necessary libraries
library(forecast)
library(quantmod)
library(ggplot2)
library(tseries)
# Define the coefficients for the MA(2) process
theta1 <- -2
theta2 <- 1.35

# Theoretical Autocorrelation for an MA(2) process
rho <- c(1,
         (theta1 + theta1 * theta2) / (1 + theta1^2 + theta2^2),
         (theta2) / (1 + theta1^2 + theta2^2),
         rep(0, 8)) # Autocorrelations for lags > 2 are 0

# Print the theoretical autocorrelation
print(rho)

# Set seed for reproducibility
set.seed(0)

# Simulate the MA(2) process
t_max <- 100
epsilon <- rnorm(t_max + 2)  # Generate white noise terms
y <- rep(0, t_max)

# Generate the time series according to the MA(2) formula
for (t in 3:(t_max + 2)) {
  y[t - 2] <- 0.7 - 2 * epsilon[t - 1] + 1.35 * epsilon[t - 2] + epsilon[t]
}

# Compute the sample autocorrelation function
sample_acf <- acf(y, lag.max=10, plot=FALSE)

# Print the sample autocorrelation function
print(sample_acf$acf)

# You can also plot the ACF for visual analysis
acf(y, lag.max=10, main="Sample Autocorrelation Function")

# To compare the theoretical and sample ACF visually, you can plot them together
# Note: You would need to adjust the indices and scaling to make the comparison.
```


```{r, error=TRUE}
#5

# Assuming y is the time series data from the previous exercise

# a. Estimate an MA(2) process
fit <- Arima(y, order=c(0,0,2))
summary(fit)

# b. Compute the forecasts
# Since we know the last two errors, we can use them for forecasting
# The 1-step ahead forecast will use both errors
# The 2-step and 3-step ahead forecasts will be the mean of the process for an MA(2)

last_error <- c(0.4, -1.2)  # Known last errors
forecasts <- numeric(3)
forecasts[1] <- fit$coef[1] * last_error[1] + fit$coef[2] * last_error[2] + mean(fit$residuals)
forecasts[2] <- mean(fit$residuals)  # 2-step ahead forecast is the mean for MA(2)
forecasts[3] <- mean(fit$residuals)  # 3-step ahead forecast is the same

# Print the forecasts
forecasts
```


```{r, error=TRUE}
#6


# Assuming that you have a time series 'y' that follows the MA(1) process with theta = 0.8
# First, simulate an MA(1) process with theta = 0.8
set.seed(123) # for reproducibility
epsilon <- rnorm(1000) # white noise
y <- filter(epsilon, sides=1, filter=c(0.8,1)) # MA(1) process

# Fit the MA(1) model to the simulated data
ma_fit <- Arima(y, order=c(0,0,1), include.mean=TRUE)

# Print the summary of the fitted model
summary(ma_fit)

# Forecast the next 10 values
forecasts <- forecast(ma_fit, h=10)

# Print the forecasts
print(forecasts)

# Plot the forecasts
plot(forecasts)
```


```{r, error=TRUE}
#10


# Download stock prices
getSymbols("AAPL", from = "2020-01-01", to = "2024-01-01")
aapl_returns <- dailyReturn(AAPL)

# Obtain autocorrelation function
aapl_acf <- acf(aapl_returns, lag.max = 20, plot = TRUE)

# Fit ARIMA model
aapl_fit <- auto.arima(aapl_returns)

# Print summary
summary(aapl_fit)

# Forecast the next 10 periods
aapl_forecasts <- forecast(aapl_fit, h = 10)

# Print the forecasts
print(aapl_forecasts)

# Plot the forecasts using autoplot
autoplot(aapl_forecasts)
```


```{r, error=TRUE}
#7.2

# Load the data
V1 <- read.table("C:/Users/valen/Downloads/labordata.dat", header = FALSE)

# Replace "path_to_your_data/labordata.dat" with the actual path to your .dat file
# Set the 'header' argument to TRUE if the first row contains column names
# Set the 'sep' argument to the character that delimits your data, such as "," for CSV files
# Add any other arguments that are necessary for your specific data format

# Inspect the data
head(V1)
summary(V1)
str(V1)

# If there are any preprocessing steps needed, perform them here


# Calculate the autocorrelation function for different displacements of time
acf(V1, lag.max=20, main="ACF for Unemployed Persons")

# Calculate the partial autocorrelation function
pacf(V1, lag.max=20, main="PACF for Unemployed Persons")

# If an AR model seems appropriate,  fit it using the `ar` function
# Assuming V1 is a data frame and you want to use the first column for the AR model
ar_model <- ar(V1[,1], method="mle") # Replace 1 with the index of the column you want to use
summary(ar_model)

# Plot the data with the AR model fit
plot(V1)
lines(fitted(ar_model), col="blue")
```

```{r, error=TRUE}
#7.5


# Simulate two stationary AR(2) processes
set.seed(123)  # Reproducibility
sim1 <- arima.sim(model = list(ar = c(0.5, -0.3)), n = 1000)
sim2 <- arima.sim(model = list(ar = c(0.4, -0.2)), n = 1000)

# Plot the time series
par(mfrow = c(2, 1))
plot.ts(sim1, main = "AR(2) Process 1: Yt = 0.5*Yt-1 - 0.3*Yt-2 + et")
plot.ts(sim2, main = "AR(2) Process 2: Yt = 0.4*Yt-1 - 0.2*Yt-2 + et")

# Compute and plot autocorrelation functions
acf(sim1, main = "ACF for AR(2) Process 1")
acf(sim2, main = "ACF for AR(2) Process 2")

# Check for stationarity using Augmented Dickey-Fuller Test
print(adf.test(sim1))
print(adf.test(sim2))

# Check for stationarity using KPSS Test
print(kpss.test(sim1))
print(kpss.test(sim2))

arima_model1 <- arima(sim1, order = c(2, 0, 0))
arima_model2 <- arima(sim2, order = c(2, 0, 0))

print(summary(arima_model1))
print(summary(arima_model2))

pacf(sim1, main = "PACF for AR(2) Process 1")
pacf(sim2, main = "PACF for AR(2) Process 2")

```


```{r, error=TRUE}
#7.6

# Simulated time series data for CPI
set.seed(123)
cpi_general <- ts(rnorm(100, mean = 2, sd = 0.5), start = c(2000, 1), frequency = 12)
cpi_excl <- ts(rnorm(100, mean = 1.8, sd = 0.6), start = c(2000, 1), frequency = 12)

# Plot both time series for visual comparison
ts.plot(cpi_general, cpi_excl, col = c("blue", "red"), lty = 1:2)

# Fit ARIMA models to both time series
fit_general <- auto.arima(cpi_general)
fit_excl <- auto.arima(cpi_excl)

# Summary of fit to check for goodness of fit
summary(fit_general)
summary(fit_excl)

# Perform forecasts
forecast_general <- forecast(fit_general, h = 2)
forecast_excl <- forecast(fit_excl, h = 2)

# Plot the forecasts with density
plot(forecast_general, main = "1-step and 2-step Density Forecast for General CPI")
plot(forecast_excl, main = "1-step and 2-step Density Forecast for CPI Excluding Gas and Food")

# Print the forecasts
print(forecast_general)
print(forecast_excl)
```


```{r}
# Perform forecasts
forecast_general <- forecast(fit_general, h = 2)
forecast_excl <- forecast(fit_excl, h = 2)

# Plot the forecasts with density
plot(forecast_general, main = "1-step and 2-step Density Forecast for General CPI")
plot(forecast_excl, main = "1-step and 2-step Density Forecast for CPI Excluding Gas and Food")

# Print the forecasts
print(forecast_general)
print(forecast_excl)
```
# Insights and Observations

## 1. Time Series Comparison

# Two time series, `cpi_general` and `cpi_excl`, were plotted.

### Visual Observations:
# - `cpi_general` is shown as a solid blue line and `cpi_excl` as a dashed red line.
# - Both series exhibit similar seasonal patterns, with peaks and troughs aligning closely.
# - `cpi_excl` has slightly higher volatility compared to `cpi_general`.


## 2. Density Forecasts

### a. General CPI
# A 1-step and 2-step ahead density forecast was generated for `cpi_general`.

#### Observations:
# - Forecasts closely follow the most recent values of the series.
# - Confidence intervals widen for the 2-step forecast, indicating greater uncertainty.

### b. CPI Excluding Gas and Food
# Similar forecasts were produced for `cpi_excl`.

#### Observations:
# - Forecasts show wider confidence intervals compared to `cpi_general`, consistent with its higher observed volatility.



## 3. Observations on Model Fit

### General CPI Model:
# - `fit_general` used an ARIMA(0,0,0)(1,0,0)[12] model.
# - AIC: 127.8, BIC: 135.62
# - Residuals suggest a reasonable fit with no strong autocorrelations.

### CPI Excluding Gas and Food Model:
# - `fit_excl` used an ARIMA(0,0,0)(0,0,1)[12] model.
# - AIC: 177.01, BIC: 184.83
# - Residuals suggest a good fit, though with slightly wider variance.



