---
title: 'Project #1'
author: "Gavin Valenzuela"
date: "2024-07-04"
output: pdf_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(width = 80)
load("C:/Users/valen/Downloads/hp.RData")
library(magrittr)
library(MASS)
library(car)
library(RRPP)
library(lmtest)

model <- lm(sell ~ . , data = hp)
```

```{r}
viftest <- vif(model)
viftest
```
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(width = 80)
cat("Removing the stories and bedrooms variables because they have the highest VIFs and it also\n")
cat("makes logical sense since the larger the lot size, the more likely more bedrooms are, so\n")
cat("lot size covers that aspect.\n")
model1 <- lm(sell ~ . - stories - bedrooms, data = hp)
#Found a command for generating best variables to use via AIC test which turns out to be all the variables.
lm(sell ~ . , data = hp) %>%
stepAIC(direction = "backward", trace = 0) %>%
summary()
```
```{r}
AIC(model)
```

```{r}
AIC(model1)
```

```{r}
# Assuming model is already defined and fitted to your data

# Generate fitted values from the model
fittedmodel <- fitted.values(model)

# Additional regressor like squared and cubed terms
fittedsquared <- fittedmodel^2
fittedcubed <- fittedmodel^3

# Make a new model with the new variables added
newmodel <- lm(sell ~ fittedsquared + fittedcubed, data = hp)

# Check the summary of the new model
summary(newmodel)
```

```{r}
#performing reset test on the new model
Rtest <- resettest(model, power = 2:3,type = "regressor")
print(Rtest)
```
#3
```{r}
# Fit the baseline model
baseline_model <- lm(sell ~ ., data = hp)

# Create a summary of the baseline model
baseline_summary <- summary(baseline_model)

# Print the summary
print(baseline_summary)
```
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(width = 80)
# Interpretation
#(Comment on the Statistical and 
#Economic Significance of Your Individual Estimates
#From the summary output of the baseline model
#*Statistically Significant Variables:
#Variables with p-values less than 0.05 are statistically significant
#Significant variables include lotsize, fullbath, stories,
#driveway, recreation, basement, gas, centralair, garage, 
#and neighbour
#*Insignificant Variables
#The variable bedrooms has a p-value of 0.0807,
#which is greater than 0.05, indicating it 
#is not statistically significant at the 5% level
```


```{r}
# Identify statistically significant variables
coef_summary <- baseline_summary$coefficients
significant_vars <-
coef_summary[coef_summary[, 4] < 0.05, ]

# Print significant variables
print(significant_vars)

```
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(width = 80)
# Economic Significance
# Interpretation of Coefficients:
# - **Lotsize:** The coefficient is 3.5463, meaning a one-unit increase in lot size is associated 
#with an increase of approximately 3.55 units in sell, holding other factors constant. 
#This is economically significant as it shows a positive impact on property value.
# - **Fullbath:** The coefficient is 14335.5585, 
#meaning an additional full bathroom is 
#associated with an increase of approximately 
#14335.56 units in sell. 
#This is a substantial impact and economically significant.
# - **Stories:** The coefficient is 6556.9457,
#meaning an additional story in the building is associated 
#with an increase of approximately 6556.95 
#units in sell.
# - **Driveway:** The coefficient is 6687.7789, 
#meaning having a driveway is associated 
#with an increase of approximately 
#6687.78 units in sell.
# - **Recreation:** The coefficient is 
#4511.2884, meaning the presence of recreation
#facilities is associated with
#an increase of approximately 4511.29 units in sell.
# - **Basement:** The coefficient is 5452.3835,
#meaning having a basement is associated 
#with an increase of approximately 5452.38 
#units in sell.
# - **Gas:** The coefficient is 12381.4066, 
#meaning the presence of a gas system 
#is associated with an increase
#of approximately 12381.41 units in sell.
# - **Centralair:** The coefficient 
#is 12632.8905, meaning the presence 
#of central air conditioning is associated
#with an increase of approximately 12632.89 units in sell.
# - **Garage:** The coefficient is 4244.8290, 
#meaning the presence of a garage is 
#associated with an increase of 
#approximately 4244.83 units in sell.
# - **Neighbour:** The coefficient is 9369.5132, 
#meaning being in a certain neighborhood 
#is associated with an increase of 
#approximately 9369.51 units in sell.
```

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(width = 80)
# - The intercept has a negative coefficient (-4038.3540), 
#which might be unexpected in some contexts 
#but can be interpreted as the base value
#when all other variables are zero.
# - The large coefficients for fullbath, gas, and centralair 
#suggest these variables have a 
#substantial economic impact, which should be verified for practical realism.

```


```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(width = 80)
# Overall fit metrics
r_squared <- baseline_summary$r.squared
adjusted_r_squared <- baseline_summary$adj.r.squared
f_statistic <- baseline_summary$fstatistic

```

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(width = 80)

#R-squared: 0.6731, indicating that approximately 
#67.31% of the variance in sell 
#is explained by the mode
#Adjusted R-squared: 0.6664
#slightly lower than R-squared, accounting for the number of predictors. 
#This is still a good fit, showing the model 
#explains a substantial portion of the variance.",
#The F-statistic is 99.97 with a p-value
#< 2.2e-16, indicating that the model is statistically
#significant overall.
#This suggests that the predictors, 
#taken together, significantly explain the variability in sell."
```


```{r}
# Print the overall fit metrics
cat("R-squared:", r_squared, "\n")
cat("Adjusted R-squared:", adjusted_r_squared, "\n")
cat("F-statistic:", f_statistic[1], "on", f_statistic[2], "and", f_statistic[3], "DF, p-value:", pf(f_statistic[1], f_statistic[2], f_statistic[3], lower.tail = FALSE), "\n")
```


```{r}

# Perform RESET test on the new model
Rtest <- resettest(baseline_model, power = 2:3, type = "regressor")
print(Rtest)
```
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(width = 80)
#The RESET test has a p-value of 0.6546, 
#which is not significant. This indicates 
#that there is no strong evidence against 
#the model specification. In other words,
#the model does not suffer from omitted
#variable bias or incorrect functional form.
# Summary
#Statistical Significance:**
#Most variables are statistically significant, #indicating they have a meaningful relationship
#with the dependent variable.
#Economic Significance:** Variables like lotsize,
#fullbath, and centralair show substantial economic
#impacts, which align with practical expectations.",
#"- **Model Fit:** The model has a good fit with 
#an R-squared of 67.31% and an
#adjusted R-squared of 66.64%. The model is also
#statistically significant 
#as indicated by the F-statistic.",
# **Model Specification** The RESET test suggests
# that the model does not have specification errors.",
#"Overall, the baseline model appears to be robust 
#and provides meaningful insights into the factors
#affecting property values (sell)."
```
#8
```{r}
library(lmtest)
library(sandwich)
```

```{r}
# Fit the baseline model
baseline_model <- lm(sell ~ ., data = hp)

# Perform the Breusch-Pagan test for heteroskedasticity
bp_test <- bptest(baseline_model)

# Print the results of the Breusch-Pagan test
print(bp_test)
```
#The Breusch-Pagan test results show 
#a BP = 61.953 with a p-value of 4.014e-09.
#Since the p-value is significantly 
#less than 0.05, we reject the 
#null hypothesis of homoscedasticity.
#This indicates the presence of
#heteroskedasticity in the model.

```{r}
 # Refit the model using robust standard errors
  robust_model <- coeftest(baseline_model, vcov = vcovHC(baseline_model, type = "HC1"))
  
  # Print the summary of the model with robust standard errors
  print(robust_model)
```
#lotsize:

#Coefficient: 3.5463
#Statistically significant with p-value < 2.2e-16
#conomically significant: A one-unit increase in lot #size is associated with an increase of approximately #3.55 units in property value (sell).
#Bedrooms:

#Coefficient: 1832.00347
#Statistically insignificant with p-value = 0.07819
#This indicates that the number of bedrooms does not #have a significant effect on property value at the 5% #significance level.
#Fullbath:

#Coefficient: 14335.55847
#Statistically significant with p-value < 2e-16
#Economically significant: An additional full bathroom #is associated with an increase of approximately #14335.56 units in property value.
#Stories:

#Coefficient: 6556.94571
#Statistically significant with p-value = 2.03e-13
#Economically significant: An additional story in the #building is associated with an increase of #approximately 6556.95 units in property value.
#Driveway:

#Coefficient: 6687.77889
#Statistically significant with p-value = 6.259e-05
#Economically significant: Having a driveway is #associated with an increase of approximately 6687.78 #units in property value.
#Recreation:

#Coefficient: 4511.28834
#Statistically significant with p-value = 0.035869
#Economically significant: The presence of recreation #facilities is associated with an increase of #approximately 4511.29 units in property value.
#Basement:

#Coefficient: 5452.38346
#Statistically significant with p-value = 0.002001
#Economically significant: Having a basement is #associated with an increase of approximately 5452.38 #units in property value.
#Gas:

#Coefficient: 12381.40627
#Statistically significant with p-value = 0.002613
#Economically significant: The presence of a gas system #is associated with an increase of approximately #12381.41 units in property value.
#Centralair:

#Coefficient: 12632.89004
#Statistically significant with p-value = 1.52e-13
#Economically significant: The presence of central air #conditioning is associated with an increase of #approximately 12632.89 units in property value.
#Garage:

#Coefficient: 4244.82900
#Statistically significant with p-value = 8.901e-06
#Economically significant: The presence of a garage is #associated with an increase of approximately 4244.83 #units in property value.
#Neighbour:

#Coefficient: 9369.51324
#Statistically significant with p-value = 7.482e-07
#Economically significant: Being in a certain #neighborhood is associated with an increase of #approximately 9369.51 units in property value.
#Overall Model Fit
#R-squared: 0.6731
#Indicates that approximately 67.31% of the variance in #sell is explained by the model.
#Adjusted R-squared: 0.6664
#Slightly lower than R-squared, accounting for the #number #of predictors. Still a good fit, showing the #model explains a substantial portion of the variance.
#F-statistic: 99.97 with a p-value < 2.2e-16
#Indicates that the model is statistically significant #overall, suggesting that the predictors, taken #together, significantly explain the variability in #sell.
#Summary
#Statistical Significance: Most variables are #statistically significant, indicating they have a #meaningful relationship with the dependent variable.
#Economic Significance: Variables like lotsize, #fullbath, and centralair show substantial economic #impacts, which align with practical expectations.
#Model Fit: The model has a good fit with an R-squared #of 67.31% and an adjusted R-squared of 66.64%. The #model is also statistically significant as indicated by #the F-statistic.
#Model Specification: The RESET test suggests that the #model does not have specification errors.
#Heteroskedasticity: The Breusch-Pagan test indicates #the presence of heteroskedasticity, which has been #corrected using robust standard errors. The corrected #model provides more reliable standard errors and #significance tests.
